### 2022-06-14 docs/index.Rmd
moving slides to bonus

---
title: "Accelerated oblique random survival forests"
author: "Byron C. Jaeger"
date: Last updated `r Sys.Date()`
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dpi = 600,
  fig.align='center',
  fig.height=5
)

library(tidyverse)
library(ggforce)
library(rpart)
library(rpart.plot)
library(party)
library(parttree)
library(palmerpenguins)
library(gt)
library(table.glue)

tar_load(
  names = c(
    bm_pred_model_viz,
    bm_pred_time_viz,
    bm_pred_viz,
    data_key
  )
)

```

```{css texts}

.medium { font-size: 25px; }
.mediumer { font-size: 40px; }
.large { font-size: 80px; }
.larger { font-size: 160px; }
.huge { font-size: 240px; }

.rainbow-text {
  background-image: linear-gradient(to left, violet, indigo, blue, green, yellow, orange, red);   -webkit-background-clip: text;
  color: transparent;
}

```

## Overview

- Random forests (axis based and oblique)

    + Decision trees

    + Random survival forests (RSF)

- Accelerating the oblique RSF

    + Newton Raphson scoring

- Benchmark

    + Datasets & learners

    + Evaluation

    + Results

- Software

    + R package & website

---
class: center, middle, inverse

# .larger[Random forests] 
## (axis based and oblique)

---
background-image: url(img/penguins.png)
background-size: 45%
background-position: 85% 72.5%

## Decision trees

- Frequently used in supervised learning.

- Partitions the space of predictor variables.

- Can be used for classification, regression, and survival analysis.

.pull-left[
.medium[_Demo_:] <br><br> Axis-based and oblique decision trees for classification of penguin species (chinstrap, gentoo, or adelie) based on bill and flipper length.<sup>1</sup>
]

.footnote[
<sup>1</sup>Data were collected and made available by [Dr. Kristen Gorman](https://www.uaf.edu/cfos/people/faculty/detail/kristen-gorman.php) and the [Palmer Station](https://pal.lternet.edu/), a member of the [Long Term Ecological Research Network](https://lternet.edu/).
]

---

Dimensions for Adelie, Chinstrap and Gentoo Penguins at Palmer Station

```{r fig_penguins_nopart}

penguins <- drop_na(penguins)

ggplot(data = penguins) +
  aes(x = flipper_length_mm, y = bill_length_mm, label = species) +
  geom_point(aes(color = species, shape = species),
             size = 3,
             alpha = 0.8) +
  geom_mark_ellipse(aes(color = species, fill = species), alpha = 0.075) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  scale_fill_manual(values = c("darkorange","purple","cyan4")) +
  labs(x = "\nFlipper length, mm",
       y = "Bill length, mm\n",
       color = "Penguin species",
       shape = "Penguin species") +
  coord_cartesian(ylim = c(30, 70),
                  xlim = c(170, 235)) +
  theme(panel.grid = element_blank(),
        legend.position = '')

```

---

Partition all the penguins into flipper length < 207 or ≥ 207 mm

```{r fig_penguins_part1}

mdl_tree <- rpart(
 formula = species ~ flipper_length_mm + bill_length_mm,
 data = penguins, 
 control = rpart.control(maxdepth = 1)
)

ggplot(data = penguins) +
  aes(x = flipper_length_mm, y = bill_length_mm, label = species) +
  geom_point(aes(color = species, shape = species),
             size = 3,
             alpha = 0.8) +
  geom_parttree(data = mdl_tree, aes(fill=species), alpha = 0.1) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  scale_fill_manual(values = c("darkorange","purple","cyan4")) +
  labs(x = "\nFlipper length, mm",
       y = "Bill length, mm\n",
       color = "Penguin species",
       shape = "Penguin species") +
  coord_cartesian(ylim = c(30, 70),
                  xlim = c(170, 235)) +
  scale_x_continuous(expand = c(0,0), breaks = c(180, 200, 220)) + 
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.grid = element_blank(),
        legend.position = '')

```

---

Partition penguins on the left side into into bill length < 43 or ≥ 43 mm

```{r fig_penguins_part2}

mdl_tree <- rpart(
 formula = species ~ flipper_length_mm + bill_length_mm,
 data = penguins, 
 control = rpart.control(maxdepth = 2)
)

ggplot(data = penguins) +
  aes(x = flipper_length_mm, y = bill_length_mm, label = species) +
  geom_point(aes(color = species, shape = species),
             size = 3,
             alpha = 0.8) +
  geom_parttree(data = mdl_tree, aes(fill=species), alpha = 0.1) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  scale_fill_manual(values = c("darkorange","purple","cyan4")) +
  labs(x = "\nFlipper length, mm",
       y = "Bill length, mm\n",
       color = "Penguin species",
       shape = "Penguin species") +
  coord_cartesian(ylim = c(30, 70),
                  xlim = c(170, 235)) +
  scale_x_continuous(expand = c(0,0), breaks = c(180, 200, 220)) + 
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.grid = element_blank(),
        legend.position = '')

# ---
# background-image: url(img/penguins_logo.png)
# background-position: 95% 5%
# background-size: 110px 120px
# layout: true

```

---

With oblique splits, partitions do not need to be rectangles

```{r fig_penguins_part2_oblique}

x_delta <- 206.5 - 170
y_delta <- 46 - 40

slope <- y_delta / x_delta
intercept <- 46 - 206.5 * slope

text_lower <-
  table_glue("bill length - {slope} * flipper length < {intercept}")
text_upper <-
  table_glue("bill length - {slope} * flipper length \u2265 {intercept}")

data_poly <- tribble(
  ~flipper_length_mm, ~bill_length_mm    , ~species,
  170               ,  30                , 'Adelie',
  206.5             ,  30                , 'Adelie',
  206.5             ,  46                , 'Adelie',
  170               ,  40                , 'Adelie',
  170               ,  40                , 'Chinstrap',
  206.5             ,  46                , 'Chinstrap',
  206.5             ,  70                , 'Chinstrap',
  170               ,  70                , 'Chinstrap',
  206.5             ,  30                , 'Gentoo',
  235               ,  30                , 'Gentoo',
  235               ,  70                , 'Gentoo',
  206.5             ,  70                , 'Gentoo',
)

ggplot(data = penguins) +
  aes(x = flipper_length_mm, y = bill_length_mm, label = species) +
  geom_polygon(data = data_poly, 
               aes(fill = species, group = species),
               alpha = 0.1,
               col = 'black') +
  geom_point(aes(color = species, shape = species),
             size = 3,
             alpha = 0.8) +
  # geom_abline(slope = slope, intercept = intercept, col = 'red')
  geom_text(
    data = tibble(flipper_length_mm = c(172, 172), 
                  bill_length_mm = c(68, 32), 
                  species = c(text_upper, text_lower)),
    hjust = 0
  ) + 
  theme_minimal() +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  scale_fill_manual(values = c("darkorange","purple","cyan4")) +
  labs(x = "\nFlipper length, mm",
       y = "Bill length, mm\n",
       color = "Penguin species",
       shape = "Penguin species") +
  coord_cartesian(ylim = c(30, 70),
                  xlim = c(170, 235)) +
  scale_x_continuous(expand = c(0,0), breaks = c(180, 200, 220)) + 
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.grid = element_blank(),
        legend.position = '')

```

---

## Random survival forests (RSFs)

1. Breiman developed the random forest, a large set of decision trees injected with randomness.<sup>1, 2</sup>

.footnote[<sup>1</sup>Breiman, Leo. "Bagging predictors." Machine learning 24.2 (1996): 123-140.<br/><sup>2</sup>Breiman, Leo. "Random forests." Machine learning 45.1 (2001): 5-32.]

---

## Random survival forests (RSFs)

1. Breiman developed the random forest, a large set of decision trees injected with randomness.<sup>1, 2</sup>

1. Hothorn and, separately, Ishwaran developed extensions of the random forest for survival outcomes.<sup>3, 4</sup>

Specifically,

- Hothorn developed the conditional inference forest (CIF)

- Ishwaran developed the random survival forest (RSF)

.footnote[<sup>3</sup>Hothorn, Torsten, et al. "Unbiased recursive partitioning: A conditional inference framework." Journal of Computational and Graphical statistics 15.3 (2006): 651-674.<br/><sup>4</sup> Ishwaran, Hemant, et al. "Random survival forests." Annals of Applied Statistics 2.3 (2008): 841-860.]

---

Each leaf in the RSF contains a Kaplan-Meier estimate of survival. In the CIF, weights are applied based on sample size.

```{r}

penguins_sim <- penguins %>%
  mutate(
    time_mean = if_else(
      bill_length_mm < 40,
      true = 80,
      false = 20
    ),
    time = rnorm(n = n(), mean = time_mean, sd = 20),
    time = pmax(time, 1),
    status = rbinom(n = n(), size = 1, prob = 0.75)
  )

# png(res = 300,
#     width = 6,
#     height = 3.75,
#     units = 'in',
#     filename = 'rpart_plot_surv.png')

mdl_ctree <-
  ctree(formula = Surv(time, status) ~ bill_length_mm,
        data = penguins_sim,
        controls = ctree_control(maxdepth = 1))

plot(
  mdl_ctree,
  inner_panel = node_inner(mdl_ctree, pval = FALSE)
)

```

---

## Random survival forests (RSFs)

1. Breiman developed the random forest, a large set of decision trees injected with randomness.<sup>1, 2</sup>

1. Hothorn and, separately, Ishwaran developed extensions of the random forest to engage with survival outcomes (CIF and RSF).<sup>3, 4</sup>

1. Zhou developed a rotation survival forest and Wang developed a survival forest with an extended predictor space.<sup>5, 6</sup>

Both Zhou and Wang's extensions were based on the CIF

.footnote[<sup>5</sup>Zhou L, et al. "Rotation survival forest for right censored data." PeerJ. 2015 Jun 11;3:e1009.<br/><sup>6</sup> Wang H, et al. "Random survival forest with space extensions for censored data." Artif Intell Med. 2017 Jun;79:52-61.]

---

## Random survival forests (RSFs)

1. Breiman developed the random forest, a large set of decision trees injected with randomness.<sup>1, 2</sup>

1. Hothorn and, separately, Ishwaran developed extensions of the random forest to engage with survival outcomes (CIF and RSF).<sup>3, 4</sup>

1. Zhou developed a rotation survival forest and Wang developed a survival forest with an extended predictor based on the CIF.<sup>5, 6</sup>

1. Jaeger developed the oblique RSF, which used penalized regression to find oblique splits.<sup>7</sup>

Jaeger showed in general benchmarks that the oblique RSF had higher prediction accuracy than axis-based RSFs. However, the obliqueRSF R package is hundreds of times slower than standard R packages for axis based RSFs.

.footnote[<sup>7</sup>Jaeger BC, et al. "Oblique random survival forests." The Annals of Applied Statistics 13.3 (2019): 1847-1883.]

---
class: center, middle
background-image: url("img/meme_slow_R.jpg")
background-size: contain

---
class: center, middle, inverse

# .larger[Accelerating the oblique RSF]

---

## Accelerating the oblique RSF

We identify linear combinations of predictor variables in non-leaf nodes by applying Newton Raphson scoring to the partial likelihood function of the Cox regression model: $$L(\beta) = \prod_{i=1}^m \frac{e^{x_{j(i)}^{'} \beta}}{\sum_{j \in R_i} e^{x_j^{'} \beta}}$$

- $x_i$ is a vector of predictors values.

- $R_i$ is the set of indices, $j$, with $T_j \geq t_i$ (i.e., those still at risk at time $t_i$)

- $t_1 < \, \ldots \, < t_m$ are the $m$ unique event times in the training data.

- $j(i)$ is the index of the observation for which an event occurred at time $t_i$.

---

## Newton Raphson scoring

Estimated regression coefficients $\hat{\beta}$ are updated in each step based on their first derivative, $U(\hat{\beta})$, and second derivative, $H(\hat{\beta})$:
$$\hat{\beta}^{k+1} =  \hat{\beta}^{k} + U(\hat{\beta} = \hat{\beta}^{k})\, H^{-1}(\hat{\beta} = \hat{\beta}^{k})$$

For statistical inference, iterate until a convergence threshold is met.

For identifying linear combination of predictors in the oblique RSF, .mediumer[🤷]

- `aorsf-fast` completes one iteration.

- `aorsf-cph` iterates until convergence or 15 iterations.

---
class: center, middle, inverse

# .larger[Benchmark]

---

## Learners

__Oblique RSFs__:

- _aorsf-fast_: the fast version of aorsf

- _aorsf-cph_: the less fast but still pretty fast version of aorsf

- _aorsf-random_: randomized coefficients (Breiman's idea)

- _aorsf-net_: aorsf's copy of obliqueRSF

- _obliqueRSF-net_: oblique RSF using penalized cox regression (the original)

---

## Learners

__Oblique RSFs__:

- .rainbow-text[_aorsf-fast_]: this is the only oblique RSF we show in the racing model plots.

- _aorsf-cph_: the less fast but still pretty fast version of aorsf

- _aorsf-random_: randomized coefficients (Breiman's idea)

- _aorsf-net_: aorsf's copy of obliqueRSF

- _obliqueRSF-net_: oblique RSF using penalized cox regression (the original)

---

## Learners

__Axis based RSFs__:

- _cif-standard_: standard CIF

- _cif-extension_: CIF with space extension

- _cif-rotate_: CIF with rotation

- _rsf-standard_: standard RSF

- _ranger-extratrees_: RSF with extremely randomized trees

---

## Learners

__Other__:

- _glmnet-cox_: Penalized Cox regression model

- _nn-cox_: Cox neural network with time-varying effects

- _xgboost-cox_: Boosted trees fitted to Cox log likelihood

- _xgboost-aft_: Boosted trees fitted to accelerated failure time.

---

## Data sets

A total of 23 risk prediction tasks in 16 data sets were analyzed.

- number of observations ranged from `r min(data_key$nrow)` to `r max(data_key$nrow)` (median = `r median(data_key$nrow)`)

- number of predictors ranged from `r min(data_key$ncol)` to `r max(data_key$ncol)` (median = `r median(data_key$ncol)`)

- % censored ranged from `r round(100*min(data_key$pcens))` to `r round(100*max(data_key$pcens))` (median = `r round(100*median(data_key$pcens))`)

(Full table is shown in the bonus slides)

---

## Evaluation

We measured performance of each learner with:

- index of prediction accuracy (IPA); __higher__ is 👍

- time-dependent concordance (C)-statistic; __higher__ is 👍

- total time to fit a model and compute predictions; __lower__ is 👍

---

## Evaluation

How do we estimate overall performance based on all 23 risk prediction tasks?

Step 1: Collect IPA and C-statistic values:

- For each of the 23 risk prediction problems,

    + split the corresponding data into a 50/50 train/test split
    + fit each learner to the training set
    + evaluate each learner's predictions in the testing set

--

Step 2: Fit a hierarchical Bayesian model to analyze posterior expected performance for each learner:

$$\text{metric} = \widehat{\gamma} \cdot \text{model} + (1\,|\, \text{data/run})$$

- `run` refers to the specific train/test split of `data`

- `metric` is either the IPA or the time-dependent C-statistic.

---

```{r bm_pred_viz_ibs, fig.height=9, fig.width=11}
bm_pred_viz$ibs_scaled
```

---

```{r bm_pred_viz_cstat, fig.height=9, fig.width=11}
bm_pred_viz$cstat
```

---

```{r bm_pred_model_viz_ibs, fig.height=6.9, fig.width=9}
bm_pred_model_viz$ibs_scaled
```

---

```{r bm_pred_model_viz_cstat, fig.height=6.9, fig.width=9}
bm_pred_model_viz$cstat
```

---

```{r bm_pred_time_ridges, fig.height=6.5, fig.width=8}
bm_pred_time_viz
```

---

## Using the accelerated oblique RSF

```{r out.width='100%'}

knitr::include_url("https://bcjaeger.github.io/aorsf/")

```

---

# Thank you!

---
class: center, middle

# .huge[.rainbow-text[BONUS ROUND]]

---

# Data sets

A total of 23 risk prediction tasks in 16 data sets were analyzed.

This table is continued on the next 2 slides.

```{r}

tbls <- data_key |> 
  select(label, outcome, nrow, ncol, nevent, pcens, pmiss, pctns) |> 
  mutate(
    across(starts_with('p'),  ~.x * 100),
    across(where(is.numeric), .fns = table_value),
    outcome = Hmisc::capitalize(outcome)
  ) |> 
  arrange(label) |> 
  mutate(grp = seq(n()),
         grp = case_when(
           grp > 2 * n()/3 ~ 3, 
           grp > 1 * n()/3 ~ 2,
           TRUE ~ 1
         )) |> 
  group_by(grp) |> 
  group_map(
    ~ .x |> 
      gt(rowname_col = 'label') |> 
      cols_label(
        outcome = 'Outcome',
        nrow = 'N Obs',
        ncol = 'N Predictors',
        nevent = 'N Events',
        pcens = '% Censored',
        pmiss = '% Missing',
        pctns = "% Continuous"
      ) |> 
      cols_align(align = 'center') |> 
      cols_align(align = 'left', columns = 'outcome')
  )

tbls[[1]]

```

---
## Data sets continued

```{r}

tbls[[2]]

```

---

## Data sets continued

```{r}

tbls[[3]]

```

---

## Evaluation (bonus round)

Consider a testing data set:
$$\mathcal{D}_{\text{test}} = \left\{ (T_i, \delta_i, x_{i}) \right\}_{i=1}^{N_{\text{test}}}.$$
Let $\widehat{S}(t \mid x_i)$ be the predicted probability of survival up to a given prediction horizon of $t > 0$. 
The Brier score at time $t$ is $$\widehat{\text{BS}}(t) = \frac{1}{N_{\text{test}}} \sum_{i=1}^{N_{\text{test}}} \left\{ \widehat{S}(t \mid x_i)^2 \cdot I(T_i \leq t, \delta_i = 1) \cdot \widehat{G}(T_i)^{-1} + \\ [1-\widehat{S}(t \mid x_i)]^2 \cdot I(T_i > t) \cdot \widehat{G}(t)^{-1}\right\}$$

where $\widehat{G}(\cdot)$ is the Kaplan-Meier estimate of the censoring distribution.

---

## Evaluation (bonus round)

As the Brier score is time dependent, integration over time provides a summary measure of performance over a range of plausible prediction horizons. The integrated Brier score is defined as

$$\mathcal{\widehat{BS}}(t_a, t_b) = \frac{1}{t_b - t_a}\int_{t_a}^{t_b} \widehat{\text{BS}}(t) dt.$$ 
In our results

- $t_a$ is the 25th percentile of event times

- $t_b$ is the 75th percentile of event times

---

## Evaluation (bonus round)

$\mathcal{\widehat{BS}}(t_a, t_b)$, a sum of squared prediction errors, can be scaled to produce a measure of explained residual variation (i.e., an $R^2$ statistic) by computing $$R^2 = 1 - \frac{\mathcal{\widehat{BS}}(t_a, t_b)}{\mathcal{\widehat{BS}}_0(t_a, t_b)}$$ where $\mathcal{\widehat{BS}}_0(t_a, t_b)$ is the integrated Brier score when a Kaplan-Meier estimate for survival based on the training data is used as the survival prediction function $\widehat{S}(t)$.

_Jargon_: $R^2$ = __index of prediction accuracy__ (IPA)