### 2022-08-01 README.Rmd
fig caps visible now

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 300,
  echo = FALSE,
  comment = "#>"
)

library(table.glue)

targets::tar_load(
  names = c(bm_pred_time_viz,
            bm_pred_model_viz,
            bm_vi_viz)
)

```

# aorsf-bench

<!-- badges: start -->
<!-- badges: end -->

The goal of aorsf-bench is to introduce and evaluate

1. a method to increase the computational efficiency of (i.e., accelerate) the oblique random survival forest (RSF)
1. a method to estimate importance of individual predictor variables with the oblique RSF.

The entire project is summarized in a paper: `paper/arxix/main.pdf`.

## tl;dr

In case you want a much quicker summary of the paper:

- We made oblique RSFs faster and also developed a new method to estimate importance of individual predictors with them. These methods are available in the `aorsf` R package.

- We find that the accelerated oblique RSF (`aorsf-fast` in the paper) is very fast. In fact, in a benchmark with 35 different risk prediction tasks, `aorsf-fast` was faster than all of the learners we analyzed, except for penalized Cox regression models (`glmnet-cox` in the paper). __Figure__: Distribution of time taken to fit a prediction model and compute predicted risk. The median time, in seconds, is printed and annotated for each learner by a vertical line.

```{r, fig-bm-time, fig.align='center', fig.width=8, fig.height=8}
bm_pred_time_viz$fig
```

- We find that `aorsf-fast` has the best index of prediction accuracy out of all the learners we evaluated: __Figure__: Expected differences in index of prediction accuracy between the accelerated oblique random survival forest and other learning algorithms. A region of practical equivalence is shown by purple dotted lines, and a boundary of non-zero difference is shown by an orange dotted line at the origin.

```{r, fig-bm-ibs, fig.align='center', fig.width=9, fig.height=8}
bm_pred_model_viz$fig$ibs_scaled
```

- We find that negation variable importance improves the chances of ranking a relevant variable as more important than an irrelevant variable when using an oblique RSF to estimate variable importance. __Figure__: Concordance statistic for assigning higher importance to relevant versus irrelevant variables. Text appears in rows where negation importance obtained the highest concordance, showing absolute and percent improvement over the second best technique.

```{r, fig-bm-vi, fig.align='center', fig.width=8, fig.height=10}
bm_vi_viz$fig
```

## Reproducing this paper

We use `targets` to coordinate the general benchmark experiment and simulation study that are presented in this research, and hope that our use of `targets` also makes it easier for others to engage with our code.

Our work can be reproduced, mostly. The only thing that can't be reproduced are the results based on datasets that we can't share with you. To replicate our work using publicly available data, you need to edit the vector of datasets passed into our `targets` pipeline (see `_targets.R`) so that it only contains publicly available data. Below is a code chunk showing which datasets you need to comment out:

```{r data-to-omit, echo=TRUE}

data_source = c(
  "veteran",
  "colon_recur",
  "colon_acm",
  "pbc_orsf",
  "time_to_million",
  "gbsg2",
  "peakV02",
  "flchain",
  "nafld",
  "rotterdam_recur",
  "rotterdam_acm",
  "actg_aids",
  "actg_death",
  # "guide_it_cvd",    
  # "guide_it_hfhosp", 
  "breast",
  # "sprint_cvd",
  # "sprint_acm",
  "nki",
  "lung",
  "lung_ncctg",
  "follic_death",
  "follic_relapse",
  "mgus2_death",
  "mgus2_pcm"
  # "mesa_hf",
  # "mesa_chd",
  # "mesa_stroke",
  # "mesa_death",
  # "aric_hf",
  # "aric_chd",
  # "aric_stroke",
  # "aric_death",
  # "jhs_stroke",
  # "jhs_chd"
)

```

You'll still need to create the `nafld` data by running `nafld_build()`, which is an R function included in `R/nafld_build.R` that combines multiple datasets in the `survival` package into a dataset we analyzed in the current study.

Once you've completed these steps and installed the packages listed in `packages.R`, all you need to do is run `targets::tar_make()`. However, it will take a very long time to make the pipeline as-is and I recommend setting `run_seed` to be `1:3` (see `_targets.R` to modify this value) to request only 3 replications of Monte-Carlo cross validation be completed. Also, you are free to use only a subset of the learners by commenting out the ones you don't want to run. Here is what I'd recommend if you'd like to speed up the pipeline:

```{r models-to-omit, echo=TRUE}

# slower model fitters are commented out

model_fitters <- c(
  'aorsf_fast',
  'aorsf_cph',
  # 'aorsf_random',
  # 'aorsf_net',
  # 'rotsf',
  # 'rsfse',
  # 'cif',
  'cox_net',
  # 'coxtime',
  # 'obliqueRSF',
  'xgb_cox',
  # 'xgb_aft',
  'randomForestSRC',
  'ranger'
)

```