### 2022-11-23 paper/jmlr/cover-letter/jmlr_coverletter.tex

% !TEX program = pdflatex
\input{jmlr_coverletter_preamble.tex}

\begin{document}
\hfill{\today}

Journal of Machine Learning Research

\bigskip

Dear Editors:

We are writing to submit our manuscript ``\red{Accelerated and interpretable oblique random survival forests}'' to the \emph{Journal of Machine Learning Research}.

Our manuscript studies the \red{computational efficiency and interpretability of oblique random survival forests}, which is important because \red{oblique random survival forests have high prediction accuracy but high computational overhead and few methods have been developed to interpret them}. This paper extends our previous work in \emph{Annals of Applied Statistics}, \red{Oblique Random Survival Forests} (DOI: 10.1214/19-AOAS1261), by addressing \red{the computational inefficiency and difficulty of interpreting oblique random survival forests}. We understand the number of pages in our submission is high. This is due to a large table included in our appendix that summarizes results from a general benchmark that includes 35 prediction tasks. We have included this table and other specific details in the appendix for completeness, but we acknowledge they are not critical to the paper and would be happy to reduce the total page count by removing select content in our appendices.

We suggest the following action editors and referees for our submission.

Action Editors:
\begin{itemize}
    \item \red{Miguel Carreira-Perpinan, University of California (mcarreira-perpinan@ucmerced.edu)}
    \item \red{Jason Klusowski, Princeton University (jason.klusowski@princeton.edu)}
    \item \red{Amichai Painsky, Tel Aviv University (amichaip@tauex.tau.ac.il)}
\end{itemize}

Reviewers:
\begin{itemize}
    \item \red{Terry M. Therneau, Mayo Clinic (therneau@mayo.edu)}
    \item \red{Jack Dunn, Massachusetts Institute of Technology (jack.dunn.nz@gmail.com)}
    \item \red{Marvin N. Wright, University of Bremen (wright@leibniz-bips.de)}
\end{itemize}

Our submission has the following keywords: \red{Oblique}, \red{Survival}, \red{Random Forests}, \red{Fast}, \red{Variable Importance}

As the corresponding author, I confirm that none of the co-authors listed below have a conflict of interest with the action editors and referees I suggest above. Further, I confirm that all co-authors below consent to my submission of this manuscript to the \emph{Journal of Machine Learning Research}.

\bigskip

Sincerely,

\medskip

\red{Byron C. Jaeger (Wake Forest University School of Medicine)}\\
\red{Sawyer Welden (Wake Forest University School of Medicine)}\\
\red{Kristin Lenoir (Wake Forest University School of Medicine)}\\
\red{Jaime L. Speiser (Wake Forest University School of Medicine)}\\
\red{Matthew W. Segar (Texas Heart Institute)}\\
\red{Ambarish Pandey (University of Texas Southwestern Medical Center)}\\
\red{Nicholas M. Pajewski (Wake Forest University School of Medicine)}

\end{document}