### 2023-04-28 paper/jcgs/cover-letter/coverletter_main.tex
paper can be made using only public data

% !TEX program = pdflatex
\input{coverletter_preamble.tex}

\begin{document}
\hfill{\today}

Journal of Computational and Graphical Statistics

\bigskip

Dear Editors:

We are writing to submit our manuscript ``\red{Accelerated and interpretable oblique random survival forests}'' to the \emph{Journal of Computational and Graphical Statistics}.

Our manuscript studies the \red{computational efficiency and interpretability of oblique random survival forests}, which is important because \red{oblique random survival forests have high prediction accuracy but high computational overhead and few methods have been developed to interpret them}. This paper extends our previous work in \emph{Annals of Applied Statistics}, \red{Oblique Random Survival Forests} (DOI: 10.1214/19-AOAS1261), by addressing \red{the computational inefficiency and difficulty of interpreting oblique random survival forests}.

We suggest the following reviewers:
\begin{itemize}
    \item \red{Terry M. Therneau, Mayo Clinic (therneau@mayo.edu)}
    \item \red{Marvin N. Wright, University of Bremen (wright@leibniz-bips.de)}
\end{itemize}

Our submission has the following keywords: \red{Oblique}, \red{Survival}, \red{Random Forests}, \red{Fast}, \red{Variable Importance}

As the corresponding author, I confirm that none of the co-authors listed below have a conflict of interest with the action editors and referees I suggest above.

\bigskip

Sincerely,

\medskip

\red{Byron C. Jaeger (Wake Forest University School of Medicine)}\\
\red{Sawyer Welden (Wake Forest University School of Medicine)}\\
\red{Kristin Lenoir (Wake Forest University School of Medicine)}\\
\red{Jaime L. Speiser (Wake Forest University School of Medicine)}\\
\red{Matthew W. Segar (Texas Heart Institute)}\\
\red{Ambarish Pandey (University of Texas Southwestern Medical Center)}\\
\red{Nicholas M. Pajewski (Wake Forest University School of Medicine)}

\end{document}